### 2015-07-08 text/ms.Rmd
Rework intro yet again

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory.  One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used in the fields of climate and weather
science, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble methods to improve estimates of fish population
status and trajectory. We fit four models of population and exploitation
dynamics to data-limited populations and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. We build our models on a simulated dataset of 5760
fish stocks and test them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments. 
We find that ensemble and superensemble models
substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained ensembles were applied to global fish stocks, the best
ensembles still improved all dimensions of performance over the next-best
individual model. Ensemble and superensemble models have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations they are applied to.

@@

Despite considerable advancements in the systematic collection of ecological
population data, we still have little data with which to estimate population
status for most species (IUCN, FAO, amphibian citations etc.). Yet, a large
portion of these data-limited populations face human-caused pressures and may
warrant conservation action (IUCN REF). This scenario calls for methods to
assess the status of populations that rely on limited sources of data. While,
such methods are increasingly available (e.g. REF, REF), their application faces
two key challenges: (1) performance is relatively poor compared to more
data-rich forms of population modelling (e.g. REF) and (2) resource managers are
often faced with reconciling multiple potentially conflicting estimates of
population status (e.g. REF).

@@

Whereas averages or weighted averages can improve predictions, they may not
optimally leverage available data. Some models perform better than others in
certain conditions and the covariance between model predictions may contain
information that can both improve predictions and aid understanding of
individual model performance. We can exploit these characteristics by using the
predictions from a group of models as inputs into a separate statistical model.
This technique, sometimes called 'superensemble' modelling, is common in climate
and weather forecasting. The superensemble is fit to some training dataset where
outcomes are well known and used to predict on a separate dataset. For example,
@krishnamurti1999 combined predictions of wind and precipitation in Asian
monsoons via a superensemble regression fit to observed data. Their
superensemble considerably outperformed any individual prediction or weighted
average of predictions in terms of accuracy.

>>>>

Ecological resource managers are often faced with reconciling multiple potentially
conflicting estimates of population status and trajectory.  One solution is to
average these predictions in an ensemble, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used in the fields of climate and weather
science, may provide a superior solution. Superensembles use the predictions
from multiple models as covariates in a new statistical model. Here we evaluate
the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population
status and trajectory. We fit four models of population and exploitation
dynamics to data-limited populations and combine their estimates of the mean and
slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an
ensemble average and three superensembles: a linear model, a random forest, and
a boosted regression tree. We build our models on a simulated dataset of 5760
fish stocks and test them with cross-validation and a global database of 
`r ram_stocks_n` stock assessments. 
We find that ensemble methods
substantially improve estimates of population status and trajectory.
Machine-learning superensembles performed the best for population status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, correlation between predicted and true
status improved from 
`r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median
proportional error) declined from 
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the trajectory of status. When the
simulation-trained superensembles were applied to global fish stocks, the best
models still improved all dimensions of performance over the next-best
individual model. Ensemble methods have the potential to
improve estimates of population status and trajectory across many taxa beyond
fish. However, they must be carefully tested, formed from a diverse set of
accurate models, and built on a dataset sufficiently representative of the
populations they are applied to.

@@

**Note if you're editing this: I recently started reworking the paper to
incorporate 'superensembles' as a term and focus on management/conservation need
up front instead of methods. As a result, some parts (mostly the introduction)
are still in rough draft form.**

\noindent The estimation of ecological population status and trajectory are two
of the most fundamental values to quantify in ecological resource management
[e.g. @mace2008]. Status and trajectory are fundamental to everything from
global assessments of extinction risk [e.g. @stuart2004; @schipper2008;
@juan-jorda2011] to local management of hunting (REF) and fishing (REF). But
ecological resource managers are often faced with reconciling multiple uncertain
and potentially conflicting estimates of these quantities---some models may
suggest a population is increasing in abundance while other may suggest it is
declining (e.g. REF).

@@

Whereas averages or weighted averages can improve predictions, they may not
optimally leverage available data. The truth does not necessarily lie in the
middle of multiple model predictions. Some models perform better than others in
certain conditions and the covariance between model predictions may contain
information that can improve predictions. We can exploit these characteristics
by using the predictions from a group of models as inputs into a separate
statistical model. This technique, sometimes called 'superensemble' modelling,
is common in climate and weather forecasting [@krishnamurti1999]. The
superensemble is fit to some training dataset where outcomes are well known and
used to predict on a dataset of interest. For example, @krishnamurti1999
combined predictions of wind and precipitation in Asian monsoons via a
superensemble regression fit to observed data. Their superensemble considerably
outperformed any individual prediction or weighted average of predictions in
terms of accuracy.
